---
title: "Using the controlVariatesME package"
author: "Keith Barnatchez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
description: >
  This vignette provides a tutorial on how to use the controlVariatesME package.
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
options(tibble.print_min = 4, tibble.print_max = 4)
```
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

\tableofcontents

## Overview

The `controlVariatesME` package provides a means for implementing the methods outlined in (arXiv link), which extend the method of control variates originally introduced by (cite Yang and Ding) to exposure measurement error problems. Broadly, the control variates method is appropriate in settings where (1) the exposure $A$ of interest is measured with error, and (2) the researcher has access to validated exposure measurements for a subset of the full dataset. See (link to arXiv paper) for full details behind the theory of the method.


## Installation

```{r, echo=FALSE}
# devtools::install_github("keithbarnatchez/controlVariatesME")
library(AIPW)
library(controlVariatesME)
library(SuperLearner)
```

## Tutorial

We'll begin by simulating data to demonstrate the use of the `controlVariatesME` package. In this example, we'll simulate a single covariate $X$, a binary exposure $A$, and an outcome $Y$. We'll also simulate a misclassified exposure $A^*$, which is generated by flipping the true exposure with probability `misclass`. We'll simulate selection $S$ into the validation dataset completely at random, with  $\mathbb{P}(S=1)=\rho$.

```{r}
n <- 2500
rho <- 0.5
misclass <- 0.15
sl.lib <- 'SL.earth'

# Simulate data for demonstrating R package usage
X <- rnorm(n) # single covariate
Z <- rnorm(n)

A <- rbinom(n,1,prob=exp(X)/(1+exp(X))) # true exposure

missclass_idx <- rbinom(n,1,prob=misclass) # incorrect classifications
Astar <- ifelse(missclass_idx==1,1-A,A) # misclassified exposure

Y <- A + X + rnorm(n) # outcome
val_idx <- rbinom(n,1,prob=rho) # validation data

df <- data.frame(X=X,Z=Z,A=A,Astar=Astar,Y=Y, S=val_idx)
```

### 1. Implementing the control variates method

Given a dataframe, one can implement the control variates method by using either the `controlvariatesME` or `controlVariatesMEGen` function. We generally recommend the use of the `controlVariatesMEGen` function. `controlvariatesMEGen` implements the control variates method with the efficient generalizability estimators outlined in the manuscript and accomodates covariate-dependent sampling schemes, while `controlVariatesME` implements validation-data-only AIPW estimators when forming $\hat \tau_\text{val}$ and $\hat \tau_\text{val}^\text{e.p.}$ and requires the validation data to be determined completely at random. Along with accommodating more general covariate-dependent sampling schemes, `controlVariatesMEGen` will typically yield a more efficient estimator.

Both `controlVariatesME` and `controlVariatesMEGen` take the following arguments:

- `df`: a dataframe containing the data

- `X`: a character vector specifying covariate name(s)

- `Y`: the name of the column in `df` containing the outcome

- `A`: the name of the column in `df` containing the true exposure

- `Astar`: the name of the column in `df` containing the error-prone exposure measurements

- `val_idx`: the name of the column in `df` indicating selection into the validation data

- `sl.lib`: character vector of SuperLearner libraries to use for estimation of the nuisance functions

```{r}
# Implement the control variates method 
X=c('X','Z') ; Y='Y' ; A='A' ; Astar='Astar' ; val_idx='S' ; sl.lib='SL.glm'


cvme_results <- controlVariatesME(df, X=X, Y=Y, A=A, Astar=Astar, 
                                  val_idx=val_idx, sl.lib=sl.lib)

# Generalizability version
cvme_results_gen <- controlVariatesMEGen(df, X=c('X','Z'), Y='Y',
                                                   A='A', Astar='Astar', 
                                              val_idx='S', sl.lib=sl.lib)
```

### 2. Extracting output 

Relevant output is stored directly in the `cvme_results` object. Point estimates and standard errors can be obtained with the `ate_hat` and `var_hat` fields. Additional stored objects include all of the objects used to construct the control variate: 

- `ate_hat_val` = $\hat{\tau}_{\text{val}}$
- `ate_hat_val_ep` = $\hat{\tau}_\text{val}^\text{e.p.}$
- `ate_hat_main_ep` = $\hat{\tau}_\text{main}^\text{e.p.}$
- `v_hat` = $\hat{v}$
- `V_hat` = $\hat{V}$
- `gamma_hat` = $\hat{\Gamma}$

